---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
activityData <- read.csv("activity.csv", stringsAsFactors = FALSE)
```


## What is mean total number of steps taken per day?
```{r}
stepsByDay <- aggregate(steps ~ date, data = activityData, FUN = sum)
stepsByDay$day <- 1:nrow(stepsByDay)

library(ggplot2)
g <- ggplot(stepsByDay, aes(day))
g + geom_bar(aes(weight=steps)) +
  labs(x = "Day Number", y = "Number of Steps")
```

Mean number of steps taken per day 
```{r}
mean(stepsByDay$steps)
```

Median number of steps taken per day 
```{r}
median(stepsByDay$steps)
```

## What is the average daily activity pattern?
```{r}
avgDailyActivity <- aggregate(steps ~ interval, data = activityData, FUN = mean)

g <- ggplot(avgDailyActivity, aes(interval, steps))
g + geom_line()
```

Calculate the interval with the maximum average number of steps 
```{r}
avgDailyActivity[avgDailyActivity$steps == max(avgDailyActivity$steps), "interval"]
```

## Imputing missing values
Calculate the total number of missing values in the dataset: 
```{r}
naData <- activityData[is.na(activityData$steps), ]
nrow(naData)
```

Fill in the missing values in the data set. Use the average number of steps for each interval:
```{r}
for (i in 1:nrow(activityData)) {
  if ( is.na(activityData[i, "steps"]) ) {
    targetInt <- activityData[i, "interval"]
    activityData[i, "steps"] <- avgDailyActivity[avgDailyActivity$interval == targetInt, "steps"]
  }
} 

head(activityData)
```

Plot the new data:
```{r}
stepsByDay <- aggregate(steps ~ date, data = activityData, FUN = sum)
stepsByDay$day <- 1:nrow(stepsByDay)

g <- ggplot(stepsByDay, aes(day))
g + geom_bar(aes(weight = steps)) + labs(x = "Day Number", y = "Number of Steps")
```

Calculate the new mean:
```{r}
mean(stepsByDay$steps)
```

```{r}
median(stepsByDay$steps)
```
## Are there differences in activity patterns between weekdays and weekends?
Convert dates to Date object 
```{r}
activityData$date <- as.Date(activityData$date)
```

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
activityData$dayType <- factor(weekdays(activityData$date) %in% c("Saturday", "Sunday"),
                               labels=c("Weekday", "Weekend"))

str(activityData)
```

Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
aggData <- aggregate(steps ~ interval + dayType, data = activityData, FUN = mean)

g <- ggplot(aggData, aes(interval, steps))
g + geom_line() + facet_grid(dayType ~ .)
```